---
title: "The Armed Conflict in Colombia"
author: "Manuela Szinovatz"
output: html_document
---

The following analysis is based on data from the Colombian governmental organisation Unidad Victimas, that is in charge of registering the victims of Columbia's Armed Conflict. The data can be downloaded here: https://cifras.unidadvictimas.gov.co/

Up until October 2018 a total number of 8.760.290 victims have been registered. Check out the table that lists the total number of violent incidents per year and find out which municipalities have been hit the hardest and when the conflict became particularly violent. Incidents reach from homicides, threats, torture, terroristic attacks and other kinds of violence. The grand majority of incidents are forced displacements.

```{r loading, warning=F, message=F, echo=F}

# load packages
library(readr)
library(tidyr)
library(dplyr)
library(readxl)
library(stringr)
library(DT)

Siniestros <- read_excel("/Users/Manu/Desktop/Kolumbien_Projekt/Medellin_R/data/raw_data_conflicto/Reporte_Siniestros.xlsx") %>%
  select(TIPO= `TIPO DE VICTIMA`,
         ANIO=`ANIO OCURRENCIA`,
         MUNICIPIO=`MUNICIPIO OCURRENCIA`,
         TOTAL)

```

```{r Data, warning=F, message=F, echo=F}

Siniestros_grouped <- Siniestros %>% 
  filter(ANIO != 9999 & ANIO >= 1998) %>% 
  group_by(MUNICIPIO, ANIO) %>% 
  summarise(Siniestros=sum(TOTAL)) %>% 
  rename(Municipality = MUNICIPIO,
         Year = ANIO,
         Incidents = Siniestros) %>% 
  arrange(desc(Incidents))

# Maybe: Siniestros_grouped$Incidents <- format(Siniestros_grouped$Incidents, big.mark = ",")

Siniestros_grouped$Municipality <- str_to_lower(Siniestros_grouped$Municipality)

Siniestros_grouped$Municipality <- str_to_title(Siniestros_grouped$Municipality)

datatable(Siniestros_grouped)

```

## Violent Incidents

The following chart shows the overall number of incidents in Colombia's most affected municipalities.

```{r visualise, echo=FALSE}

# Bringing in the packages
library(ggplot2)
library(ggthemes)
library(forcats)

#Preparing data for visualisation
Siniestros_vis <- Siniestros %>%
  filter(MUNICIPIO!="Sin Información" & ANIO != 9999 & ANIO >= 1998) %>% 
  group_by(MUNICIPIO) %>% 
  mutate(Siniestros_Total=sum(TOTAL)) %>% 
  mutate(SINIESTROS=round(Siniestros_Total/1000,1)) %>% 
  filter(Siniestros_Total>=50811)

#Plot it out
ggplot(Siniestros_vis,
       aes(x=SINIESTROS,
           y=fct_reorder(MUNICIPIO, SINIESTROS, desc=TRUE))) +
  geom_segment(
    aes(x = 0,
        xend = SINIESTROS,
        yend = fct_reorder(MUNICIPIO, SINIESTROS, desc=TRUE)),
    color = "gray50") +
  geom_point(color="red")+
  scale_x_continuous(limits=c(0,350))+
  theme_fivethirtyeight()+
  labs(x="Siniestros in Thousands",
       title="Municipalities with most Incidents",
       subtitle="Number of Violent Incidents from 1985-2018 in Thousands",
       caption="Source: https://cifras.unidadvictimas.gov.co/")+
  geom_text(aes(label=SINIESTROS), hjust=-.5) +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_blank(),
        axis.text.x = element_blank())

```

Buenaventura counts more than 40 percent more violent incidents than the second-most affected municipality Medellín, although Medellín is 6.5 times bigger than Buenaventura in terms of population.

The following chart shows how the amount of violent incidents in Buenaventura and Medellín have development over the last 33 years of conflict.

```{r Timeline, echo=FALSE, warning=F, message=F}

Siniestros_1 <- Siniestros %>%
  group_by(ANIO, MUNICIPIO) %>% 
  filter(MUNICIPIO=="MEDELLIN" | MUNICIPIO=="BUENAVENTURA") %>% 
  filter(ANIO>=1985) %>% 
  summarise(Siniestros=sum(TOTAL))

Siniestros_Col <- Siniestros %>% 
  group_by(ANIO) %>% 
  summarise(COLOMBIA_Total=sum(TOTAL))

Siniestros_Col <- Siniestros_Col %>% 
  mutate(COLOMBIA=round(COLOMBIA_Total/(nrow(Siniestros_Col)))) %>% 
  select(-COLOMBIA_Total)

Siniestros_2 <- Siniestros_1 %>%
  group_by(ANIO, MUNICIPIO) %>% 
  summarise(Siniestros_Total=sum(Siniestros)) %>% 
  spread(MUNICIPIO, Siniestros_Total)

joined_Sin <- left_join(Siniestros_2, Siniestros_Col, by="ANIO")

Timeline_Sin <- joined_Sin %>% 
  gather(MUNICIPIO, Siniestros, 2:4)

Timeline_Sin_copy <- Timeline_Sin %>% 
  filter(MUNICIPIO=="COLOMBIA") %>% 
  select(-MUNICIPIO)

Timeline_Sin <- Timeline_Sin %>% 
  filter(MUNICIPIO!="COLOMBIA")

ggplot(data=Timeline_Sin,
       aes(x=ANIO,
           y=Siniestros,
           color=MUNICIPIO)) +
  geom_line(data=Timeline_Sin_copy, color="grey")+
  geom_line()+
  scale_x_continuous(breaks=seq(1985,2015,5), limits=c(1984,2018))+
  facet_wrap(~MUNICIPIO)+
  labs(x="Year",
       title="Incidents in Buenaventura and Medellín",
       subtitle="Total Number of Incidents from 1985-2018 Compared to Colombian Average",
       caption="Source: https://cifras.unidadvictimas.gov.co/")+
  theme_fivethirtyeight()+
  theme(legend.position = "none")+
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

```
