---
title: "Social Stratification in Medellín"
author: "Manuela Szinovatz"
output: html_document
---



The following analysis is based on data from the Colombian government and was originally collected for the social program "Sisbén" (https://www.sisben.gov.co/sisben/paginas/que-es.aspx) that helps the government to allocate subsidies to people in need.

The data can be downloaded here: https://www.datos.gov.co/Salud-y-Protecci-n-Social/Base-de-Datos-Sisben-2017-Municipio-de-Medell-n/cnr8-armb

The dataset consists of 1,726,698 individual records. Each record accounts for one person living in Medellín's urban area and gives information about her/his socioeconomic situation. The data is anonymous. Although the sample is very big its accuracy varies slightly. Fewer data is available in the richer neighbourhoods, where fewer people rely on help from the social Sisbén system. This gets particularly clear looking at Comuna 14, Poblado, where out of approximately 130,000 inhabitants only 5,670 records can be found in the Sisbén dataset. In poorer districts and neighbourhoods usually between 80 and 100 percent are recorded. This leads to the assumption that the already very obvious difference between the richer and the poorer districts is even bigger in reality.

Based on that information the following datatable summarises the average Estrato as well as the average level of education and monthly income in Colombian pesos in Medellín's various neighbourhoods and the respective district (Comuna) the neighbourhood belongs to.


```{r, Import, echo=F, warning=F, message=F, include=FALSE}

# load packages
library(readr)
library(tidyr)
library(dplyr)
library(readxl)
library(stringr)
library(DT)
library(sf)
library(tigris)
library(leaflet)
library(stringr)
library(scales)
library(ggplot2)
library(ggthemes)
library(forcats)

#Import Data

Medellin <- read_csv("/Users/Manu/Desktop/Kolumbien_Projekt/Medellin_R/data/raw_data/Base_de_Datos_Sisben_2017.csv")

#Delete unneccessary columns & split in Comunas and Corregimientos

Medellin_Comunas <- Medellin %>% 
  select(COMUNA, BARRIO, NOMBARRIO, ESTRATO, VIVIENDA, TCUARTOSVI, TPERSONA,
         PARED:ACUEDUC, TENEVIV, USOSANI:SANITAR, DUCHA,
         PREPARAN:COCINAN,
         MOTO, AUTO1, 
         SEXO, ESTCIVIL, EMBARAZA, CARNET, NIVEL,
         FECHANTO,
         ACTIVI:PERCIBE) %>%
  filter(NOMBARRIO!="SANTA ELENA",  NOMBARRIO!="SAN ANTONIO DE PRADO",
         NOMBARRIO!="SAN CRISTÓBAL", NOMBARRIO!="ALTAVISTA",
         NOMBARRIO!="SAN SEBASTIAN DE PALMITAS") %>% 
  filter(COMUNA!="50", COMUNA!="60" & COMUNA!="70", COMUNA!="80", COMUNA!="90")

Total <- nrow(Medellin_Comunas)

#Calculate age from Birthday

library(lubridate)

Medellin_Comunas$NACI <- dmy(Medellin_Comunas$FECHANTO)

today <- now()
Medellin_Comunas$AGE_DAYS <- difftime(today, Medellin_Comunas$NACI)

Medellin_Comunas$AGE <- as.numeric(Medellin_Comunas$AGE_DAYS)/365.25

Medellin_Comunas$AGE <- floor(Medellin_Comunas$AGE)

Medellin_Comunas <- Medellin_Comunas %>% 
  select(-FECHANTO,-NACI, -AGE_DAYS)


```



```{r, Datatable, echo=F, warning=F, message=F}

# Calculate Average Education

Medellin_Average_Education_DT <- Medellin_Comunas %>% 
  select(COMUNA, NOMBARRIO, ESTRATO, SEXO, NIVEL, INGRESOS, AGE) %>% 
  group_by(NOMBARRIO, COMUNA) %>% 
  summarise(Average_Education=round(mean(NIVEL)))

Medellin_Average_Estrato_DT <- Medellin_Comunas %>% 
  select(COMUNA, NOMBARRIO, ESTRATO, SEXO, NIVEL, INGRESOS, AGE) %>% 
  group_by(NOMBARRIO, COMUNA) %>% 
  summarise(Average_Estrato=round(mean(ESTRATO)))

Medellin_Education_DT <- left_join(Medellin_Average_Estrato_DT, Medellin_Average_Education_DT)

Medellin_Average_Ingresos_DT <- Medellin_Comunas %>% 
  select(COMUNA, NOMBARRIO, ESTRATO, SEXO, NIVEL, INGRESOS, AGE) %>% 
  group_by(NOMBARRIO, COMUNA) %>% 
  summarise(Average_Ingresos=round(mean(INGRESOS)))

Medellin_Datatable_1 <- left_join(Medellin_Education_DT, Medellin_Average_Ingresos_DT)

Medellin_Datatable_1 <- Medellin_Datatable_1 %>% 
  mutate('District'=ifelse(COMUNA=="1", "Popular",
                ifelse(COMUNA=="2", "Santa Cruz",
                       ifelse(COMUNA=="3", "Manrique",
                              ifelse(COMUNA=="4", "Aranjuez",
                                     ifelse(COMUNA=="5", "Castilla",
                                            ifelse(COMUNA=="6", "Doce de Octubre",
                                                   ifelse(COMUNA=="7", "Robledo",
                                                          ifelse(COMUNA=="8", "Villa Hermosa",
                                                                 ifelse(COMUNA=="9", "Buenos Aires",
                                                                        ifelse(COMUNA=="10", "La Candelaria",
                                                                               ifelse(COMUNA=="11", "Laureles",
                                                                                      ifelse(COMUNA=="12", "La América",
                                                                                             ifelse(COMUNA=="13", "San Javier",
                                                                                                    ifelse(COMUNA=="14", "Poblado",
                                                                                                           ifelse(COMUNA=="15", "Guayabal",
                                                                                                                  ifelse(COMUNA=="16", "Belén", "")))))))))))))))))

Medellin_Datatable_1$District <- as.character(Medellin_Datatable_1$District)

Medellin_Datatable_1 <- Medellin_Datatable_1 %>%
  filter(Average_Ingresos!="0") %>% 
  select(Neighbourhood = NOMBARRIO,
         District,
         'District Number' = COMUNA,
         'Average Estrato' = Average_Estrato,
         'Average Education' = Average_Education,
         'Average Monthly Income' = Average_Ingresos)

Medellin_Datatable_1$Neighbourhood <- str_to_lower(Medellin_Datatable_1$Neighbourhood)

Medellin_Datatable_1$Neighbourhood <- str_to_title(Medellin_Datatable_1$Neighbourhood)

datatable(Medellin_Datatable_1)

```


# Moravia vs. Medellín

The following chart compares the percentage of people living in each Estrato in Moravia to Medellín's average. Both in Moravia and in Medellín most people belong to Estrato Two - "Low". However, the lower Estratos are generally overrepresented in Moravia.


```{r, Load Moravia, echo=F, warning=F, message=F}

Moravia <- Medellin_Comunas %>%
  filter(NOMBARRIO=="MORAVIA") %>%
  select(-COMUNA, -BARRIO, -NOMBARRIO)

Total_Mor <- nrow(Moravia)

Poblacion_Mor <- 37000

```


```{r, Visualise Moravia/Medellin, echo=F, warning=F, message=F}

Moravia_Estrato_grouped <- Moravia %>% 
  group_by(ESTRATO) %>%
  summarise(Population=n()) %>%
  mutate(Percent=round(Population/Total_Mor*100,2))

Moravia_Estrato_grouped <- Moravia_Estrato_grouped %>% 
  rename(Population_Mor=Population,
         Percent_Mor=Percent)

Medellin_Estrato_grouped <- Medellin_Comunas %>% 
  group_by(ESTRATO) %>%
  summarise(Population=n()) %>%
  mutate(Percent=round(Population/Total*100,2))

Estratos_Med_Mor <- full_join(Moravia_Estrato_grouped, Medellin_Estrato_grouped)

Estratos_Med_Mor_Per <- Estratos_Med_Mor %>% 
  select(ESTRATO, 
         Medellín=Percent,
         Moravia=Percent_Mor) %>% 
  gather(Place, Percentage, 2:3)

ggplot(data=Estratos_Med_Mor_Per,
       aes(x=ESTRATO, y=Percentage, fill=Place)) +
  geom_bar(position="dodge", stat="identity")+
  labs(x="Stratus",
       title="Stratification System in Medellín",
       subtitle="Percentage of Homes in Moravia compared to Medellín Average",
       caption="Source: Gobierno Digital Colombia",
       y="Percent") +
  theme_fivethirtyeight()+
  theme(panel.border = element_blank(), 
        panel.grid.minor = element_blank())+
  scale_x_continuous(breaks=seq(0,6,1))+
  scale_fill_brewer(palette="Accent")+
  theme(legend.position = "bottom")+
  theme(axis.title = element_text())+
  theme(panel.grid.major.x = element_blank())

```


## Looking at Moravia

The following chart shows the exact percentages of people living in Estrato 0 to 4. The higher Estratos 5 and 6 don't exist in Moravia.


```{r, Viz Moravia, echo=F, warning=F, message=F}

ggplot(data=Moravia_Estrato_grouped,
         aes(x=ESTRATO,
             y=Percent_Mor,
             fill=ESTRATO)) +
  geom_bar(stat="identity")+
  labs(x="Stratus",
       y= "Percent",
       title="Stratification System in Moravia",
       subtitle="Percentage of People living in Estratos 0 to 4",
       caption="Source: Gobierno Digital Colombia") +
  theme_fivethirtyeight()+
  theme(panel.border = element_blank(), 
        panel.grid.minor = element_blank())+
  theme(legend.position = "none")+
  theme(axis.title = element_text())+
  theme(panel.grid.major.x = element_blank())+
  geom_text(aes(label=Percent_Mor), vjust=-.25, fontface = "bold", size=4.5)

```

